Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2015-08-30 patchy-selection-paper.tex
.

\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\Exp}{Exp}
\DeclareMathOperator{\sn}{sn}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\one}{\mathbf{1}}
\newcommand{\calE}{\mathcal{E}}
\newcommand{\calC}{\mathcal{C}}
\newcommand{\dconv}{\xrightarrow{d}}
\newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
\newcommand{\given}{\,\vert\,}

@@

On the other hand, XXX WHAT IS $q(S)$ XXX.

Combining these, and absorbing the term $\E[W|W>1]$ from equation \eqref{eqn:mean_K} 
into the constant $C$,
\begin{align} \label{eqn:migrate}
  \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m} 
%     \begin{cases}
%         q(x) \times 4 \rho s_m s_p / \xi^2  \qquad & \text{ if } s_p < s_m \\
%         q(x) \times 2 \rho s_m^2 \qquad & \text{ if } s_p > s_m 
%     \end{cases} .
\end{align}
When we use this below, we set $p_e = 2 s_p / \xi^2$.
Note that the approximation has the undesirable property that it is
non-monotonic in $s_m$, when we expect the rate of adaptation via
migration to be a decreasing function of $s_m$. However, it only increases for small $s_m$, 
and is a decreasing function of $s_m$ for $s_m > |y|^2 / 2 \sigma^2$, 
which is the parameter region that we are restricted to by other assumptions.
\linelabel{rr:sm_condition}

@@

>>>>

\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\Exp}{Exp}
\DeclareMathOperator{\sn}{sn}
\DeclareMathOperator{\erf}{erf}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\one}{\mathbf{1}}
\newcommand{\calE}{\mathcal{E}}
\newcommand{\calC}{\mathcal{C}}
\newcommand{\dconv}{\xrightarrow{d}}
\newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
\newcommand{\given}{\,\vert\,}
\newcommand{\st}{\,\colon\,}

@@

The remaining term is $q(S)$, the integral of equation \eqref{eqn:eqfreq} over the patch.
Suppose the new patch is at distance $x$ from the original patch.
If the new patch has width smaller than $\sigma/\sqrt{2s_m}$, 
then $q(S)$ is approximately $q(x)$ multiplied by the area of the new patch.
If $S$ is larger, then we should use the area of the closest strip
of width $\sigma/\sqrt{2s_m}$ to the original patch;
see section \ref{apx:qS} for more details.
In either case, denote this area $A$.

Combining these, and absorbing the term $\E[W|W>1]$ from equation \eqref{eqn:mean_K} 
into the constant $C$,
\begin{align} \label{eqn:migrate}
  \migrate(x) &\approx C A \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m}  ,
%     \begin{cases}
%         q(x) \times 4 \rho s_m s_p / \xi^2  \qquad & \text{ if } s_p < s_m \\
%         q(x) \times 2 \rho s_m^2 \qquad & \text{ if } s_p > s_m 
%     \end{cases} .
\end{align}
where $A$ is the area of the new patch if that is no larger than $\sigma/\sqrt{2s_m}$.
When we use this below, we set $p_e = 2 s_p / \xi^2$.
Note that the approximation has the undesirable property that it is
non-monotonic in $s_m$, when we expect the rate of adaptation via
migration to be a decreasing function of $s_m$. However, it only increases for small $s_m$, 
and is a decreasing function of $s_m$ for $s_m > |y|^2 / 2 \sigma^2$, 
which is the parameter region that we are restricted to by other assumptions.
\linelabel{rr:sm_condition}

@@

%%%%%%% %%%%%%%%%%%%
\subsection[The Integral $q(S)$]{The integral $q(S)$}
\label{apx:qS}

In the developement above
we need to approximate the integral of equation \eqref{eqn:eqfreq}
across the area occupied by the new patch,
$q(S) = \int_S q(z) dz$.
The precise answer depends on the shape and orientation of the patches;
but we aim for a usable approximation,
primarily in terms of the shortest distance from the old patch to the new patch, denoted $x$,
and assuming $x$ is large enough we can take equation \eqref{eqn:eqfreq} as an equality.

Since $q(x)$ decreases as $|x|$ increases,
\begin{align}
    q(S) \le |S| q(x) % C \left( \frac{x\sqrt{2s_m}}{\sigma} \right)^{(d-1)/2} e^{-x\sqrt{2s_m}/\sigma} ,
\end{align}
where $|S|$ is the area of $S$.
This will be a good approximation if $S$ is small.

In one dimension, if $S$ has length $\ell$,
\begin{align}
    q(S) &= \int_x^{x+\ell} C e^{-u\sqrt{2s_m}/\sigma} du \\
         &= C e^{-x\sqrt{2s_m}/\sigma} \frac{\sigma}{\sqrt{2s_m}} \left( 1 - e^{-\ell\sqrt{2s_m}/\sigma} \right) \\
         &\le C e^{-x\sqrt{2s_m}/\sigma} \frac{\sigma}{\sqrt{2s_m}} .
\end{align}

In two dimensions, 
suppose that $T$ is a rectangle enclosing $S$ with one axis aligned towards the original patch and transverse width $w$.
Then
\begin{align}
    q(S) &\le \int_T q(x) dx \\
        &\le w \int_x^\infty q(r) r dr \\
        &= w C \int_x^\infty \left( \frac{r\sqrt{2s_m}}{\sigma} \right)^{-1/2} e^{-r\sqrt{2s_m}/\sigma} dr \\
        &\le w \frac{\sigma}{\sqrt{2s_m}} C \sqrt{\pi} \left( \frac{x\sqrt{2s_m}}{\sigma} \right)^{-1/2} e^{-x\sqrt{2s_m}/\sigma} dr .
\end{align}
If we absorb $\sqrt{\pi}$ into the constant $C$,
this is of the form $(\text{width}) \times (\sigma/\sqrt{2s_m}) \times q(x)$,
i.e., roughly the area, after replacing the length by $\sigma/\sqrt{2s_m}$.

In both cases,
the upper bound is $q(S) \le A q(x)$,
where $A$ is the area of the parts of $S$ that are no more than $x+\sigma/\sqrt{2s_m}$ away from the original patch.
Lower bounds could be obtained,
but we stick with this rough approximation.
